---
title: "V2 TabularData Bivariate Class Exercises"
author: "Vitalii Zakhozhyi"
---

**Calling data and preparing it for exercises**

```{r collect, eval=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)  
library(magrittr)
library(tibble)
library(ggrepel)

link="https://github.com/EvansDataScience/data/raw/master/crime.RData"
load(file = url(link))
```


**Exercise 1: Turn the bars into lollipop with the right components.**

Excercise 1 solution:

Creating the dataframe for what we going to plot

```{r table, eval=TRUE}
CrimeCol=table(crime$crimecat,crime$Occurred.DayTime)%>%
         prop.table(margin = 2)%>%  
         "*"(100)%>%
         round(3)

df.C=as.data.frame(CrimeCol)
colnames(df.C)=c('Crime','Daytime','Percent')
```

Building a plot using ggplot:

```{r lollipopcrime, eval=TRUE}
base = ggplot(df.C, aes(x = reorder(Crime, Percent), y = Percent))
lol1 = base + geom_segment(aes(y = 0,
                           x = reorder(Crime, Percent),
                           yend = Percent,
                           xend = reorder(Crime, Percent))) +
              geom_point()
lol2 = lol1 + facet_wrap( ~ Daytime ,nrow = 1) 
lol3 = lol2 + coord_flip() + 
              scale_y_continuous(labels=scales::unit_format(suffix = '%'))
lol4 = lol3 + labs(title='Crime frequencies in Seattle during different daytime',
                   x ='Type of crime',
                   y = 'Percent',
                   caption = 'Source: Seattle Open Data Portal')
lol5 = lol4 + theme(axis.text.y = element_text(size=5, angle = 45),
                    axis.text.x = element_text(size=7),
                    plot.title = element_text(size = 15, hjust = 0.5),
                    plot.caption = element_text(hjust = 1, vjust = -1.2),
                    axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.75),
                    axis.title.x = element_text(size = 12, hjust = 0.5, vjust = -1.25))
lol5
```


**Exercise 2: Change the heatplot to ascending order, where intensity goes from yellow to purple.**

Excercise 2 solution:

```{r heatDAscending, eval=TRUE}
base  = ggplot(df.C, aes(x = Daytime, y = reorder(Crime, -Percent), fill = Percent)) #changing order to ascending
heat1 = base + geom_tile() 
heat2 = heat1 + scale_fill_gradient(low = "yellow", 
                                   high = "purple")
heat3 = heat2 + theme_classic() 
heat4 = heat3 + theme(axis.text.x = element_text(size=10, vjust = 0.6),
                      axis.text.y = element_text(size=6, angle = 30),
                      plot.title = element_text(size = 15, hjust = 0.5),
                      plot.caption = element_text(hjust = 1, vjust = -1.2),
                      axis.title.y = element_text(size = 12, hjust = 0.5, vjust = 1.75),
                      axis.title.x = element_text(size = 12, hjust = 0.5, vjust = -1.25),
                      legend.title = element_blank(),
                      legend.position="top", 
                      legend.direction="horizontal",
                      legend.key.width=unit(1, "cm"),
                      legend.key.height=unit(1, "cm")) 
heat5 = heat4 + labs(y="Crime",
                     title='Crime frequencies in Seattle during different daytime',
                     caption = 'Source: Seattle Open Data Portal')
heat5
```


**Exercise 3: Complete the information needed in the previous plots.**

Excercise 3 solution:

Creating the dataframe for what we going to plot

```{r, eval=TRUE}
crimeY2000=crime[(crime$DaysToReport>=365) & (crime$year>=2000),]
crimeY2000$YearsToReport=crimeY2000$DaysToReport/365
crimeYearGG=crimeY2000[complete.cases(crimeY2000$YearsToReport),]

#same but only for crimes after 2000 and the reporting time less than a year
crimeWeek=crime[(crime$DaysToReport<365) & (crime$year>=2000),]
crimeWeek$WeeksToReport=crimeWeek$DaysToReport/7
crimeYearGG2=crimeWeek[complete.cases(crimeWeek$WeeksToReport) &complete.cases(crimeWeek$crimecat),]
```

Building plots using ggplot:

First plot shows max, min and median time to report crimes for crimes after the year of 2000, that took a year or longer to report.

```{r, eval=TRUE}
base= ggplot(crimeYearGG,
             aes(x = reorder(crimecat, YearsToReport, FUN = max),
                 y=YearsToReport)) 
mins = base + stat_summary(fun.y=min,size=1,
                           geom="line",show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median, size=1,
                                    geom="line",
                                    aes(group=1,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))

all4 = all3 + theme_minimal()+
              labs(title = "Average period to report crimes", 
                   subtitle = "(for crimes that took a year or longer to report)",
                    x ="Type of Crime", 
                    y = "Years to report",
                    caption = "Source: Seattle Open Data Portal") +
              coord_flip() +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.subtitle = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0))
              
all4
```

Second plot shows max, min and median time to report crimes for crimes after the year of 2000, that took less than a year to report.

```{r, eval=TRUE}
base = ggplot(crimeYearGG2,
              aes(x = reorder(crimecat, WeeksToReport, FUN = max),
                  y=WeeksToReport)) 
mins = base + stat_summary(fun.y=min,size=1.5,
                           geom="line",show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median,size=1,
                                    geom="line", linetype='dashed',
                                    aes(group=2,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))
all4 = all3 + theme_minimal()+
              labs(title = "Average period to report crimes", 
                   subtitle = "(for crimes that took less than a year to report)",
                    x ="Type of Crime", 
                    y = "Weeks to report",
                    caption = "Source: Seattle Open Data Portal") +
              coord_flip() +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.subtitle = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0))
              
all4
```

**Exercise 4: Complete the elements missing in the previous plots.**

Excercise 4 solution:

Preparing information for plotting

```{r, eval=TRUE}
daysByNeigh=tapply(crime$DaysToReport, crime$Neighborhood, mean,na.rm=T)

crimesByNeigh=tapply(crime$crimecat, crime$Neighborhood, length)%>%      
                     prop.table()%>%
                     "*"(100)%>% 
                     round(2)

daysByNeigh=as.data.frame(daysByNeigh)%>%rownames_to_column()
crimesByNeigh=as.data.frame(crimesByNeigh)%>%rownames_to_column()
num_num=merge(daysByNeigh,crimesByNeigh)
```

Building plots using ggplot:

```{r, eval=TRUE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh,label=rowname)) 
plot1= base +  geom_point() 
plot2=  plot1 + geom_text_repel(aes(label=ifelse(crimesByNeigh>=5,
                                           num_num$rowname, "")))

plot3= plot2 + labs(title='Are crimes in neighborhoods with high crime rate reported fast?', 
                 x='Average Days to Report', 
                 y='Share of Total Crime (%)',
                 caption='Source: Seattle Open Data Portal')+
                theme(plot.caption = element_text(hjust = 0),
                 plot.title = element_text(hjust = 0.5))
plot3
```

```{r, eval=TRUE}
base = ggplot(num_num, aes(daysByNeigh,crimesByNeigh)) 
scatp1 = base +  stat_density_2d(aes(fill = ..density..), 
                                 geom = "raster", contour = FALSE)
scatp2=scatp1+geom_text_repel(aes(label=ifelse(crimesByNeigh>=5,
                                               num_num$rowname, "")))
scatp3 = scatp2 +  theme(legend.position='none') 
scatp4= scatp3 + scale_fill_distiller(palette="Reds", direction=1) 
scatp5 = scatp4 +  scale_x_continuous(expand = c(0, 0)) + 
         scale_y_continuous(expand = c(0, 0)) 
scatp6= scatp5+ labs(title='Are crimes in neighborhoods with high crime rate reported fast?', 
                 x='Average Days to Report', 
                 y='Share of Total Crime (%)',
                 caption='Source: Seattle Open Data Portal')+
                theme(plot.caption = element_text(hjust = 0),
                 plot.title = element_text(hjust = 0.5))
scatp6
```

